---
title: "Analysis of Vancomycin Dispatch Times"
author: "Brian Gulbis"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  tint::tintHtml:
    self_contained: yes
link-citations: yes
---

```{r setup, include=FALSE}
library(tint)
# invalidate cache when the package version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tint'), echo = FALSE, warning = FALSE)
options(htmltools.dir.version = FALSE)
```

```{r, message=FALSE}
library(tidyverse)
library(stringr)
library(forcats)
library(lubridate)
library(broom)
library(plotly)
library(qicharts)

x <- dirr::get_rds("../data/tidy")

end_date <- "2017-03-05"

hvi <- c("HH CVICU", "HH CVIMU", "HH HFIC", "HH HFIM", "HH 5HVI", "HH CCU", "HVI CIMU")
# hvi <- "HH CVICU"

orders_levels <- full_join(vanc_levels, orders_valid, by = c("pie.id", "order.id")) %>%
    filter(event.unit %in% hvi | (is.na(event.unit) & order.unit %in% hvi)) %>%
    mutate(not_collected = is.na(Collected) & is.na(lab.result),
           location = coalesce(event.unit, order.unit),
           month = floor_date(detail.datetime, unit = "month"),
           week = floor_date(detail.datetime, unit = "week"),
           day = wday(detail.datetime, label = TRUE, abbr = TRUE),
           hour = hour(detail.datetime),
           day_order = wday(order.datetime, label = TRUE, abbr = TRUE),
           hour_order = hour(order.datetime),
           exceed_2hr = abs(collect_detail_diff) > 120,
           pilot = detail.datetime >= mdy("1/30/2017", tz = "US/Central"),
           future_order = order_detail_diff > 1) %>%
    dmap_at(c("timely120", "timely60", "appropriate"), ~ coalesce(.x, FALSE)) %>%
    dmap_at("future_order", ~if_else(.x, "Future", "Now"))
```

```{r, eval=FALSE}
write_csv(orders_levels, "../data/external/vancomycin_orders_levels.csv")

orders_levels_summary <- select(orders_levels, pie.id, order.id, event.unit, order.unit, priority, order.datetime, detail.datetime, dispatch.datetime = Dispatched, collected.datetime = Collected, order_detail_diff, order_dispatch_diff, dispatch_detail_diff, collect_detail_minutes = collect_detail_diff, timely = timely60, shift, day, hour, pilot, future_order)

write_csv(orders_levels_summary, "../data/external/vancomycin_orders_levels_key_variables.csv")
```

```{r data_sets}
data_orders <- orders_levels %>%
    filter(priority %in% c("Routine", "Stat", "Timed Study"),
           detail.datetime >= order.datetime - hours(1)) %>%
    mutate(priority = if_else(early_am, "Early AM", priority))

data_orders_backdated <- orders_levels %>%
    filter(priority %in% c("Routine", "Stat", "Timed Study"),
           detail.datetime < order.datetime - hours(1))
```

# Distribution

In all of the following figures, values where the scheduled collection time occured > 1 hour prior to the order time were removed (n = `r nrow(data_orders_backdated)`). It is unclear exactly why the the scheduled collection time for these orders was back-dated to be prior to the time of ordering. 

```{r, fig.cap="Distribution of time from task dispatch to scheduled collection"}
plot_ly(data_orders, x = ~dispatch_detail_diff) %>%
    add_histogram(name = "All") %>%
    add_histogram(split = ~priority, visible = FALSE) %>%
    add_histogram(split = ~future_order, visible = FALSE) %>%
    layout(xaxis = list(title = "Time (Hours)"),
           updatemenus = list(list(
               y = 0.8, 
               buttons = list(
                   list(method = "restyle", args = list("visible", c(list(TRUE), as.list(rep(FALSE, times = 6)))), label = "All"),
                   list(method = "restyle", args = list("visible", c(list(FALSE), as.list(rep(TRUE, times = 4)), list(FALSE, FALSE))), label = "Priority"),
                   list(method = "restyle", args = list("visible", c(list(FALSE), as.list(rep(FALSE, times = 4)), list(TRUE, TRUE))), label = "Timing")
               ))
           ))
# tickvals = seq(-24, 48, 6)
```

It is clear that Stat orders dispatch immediately, but Routine and Timed Study are much more variable, with most dispatching 1-3 hours before the scheduled collection time.

```{r, fig.cap="Distribution of time from task dispatch to scheduled collection"}
plot_ly(data_orders, x = ~priority, y = ~dispatch_detail_diff) %>%
    add_boxplot(name = "Overall") %>%
    add_boxplot(split = ~future_order) %>%
    layout(xaxis = list(title = "Collection Priority"), 
           yaxis = list(title = "Time (Hours)"))
# tickvals = seq(-24, 48, 3)
```

```{r}
x <- data_orders %>%
    group_by(future_order) %>%
    do(tidy(summary(.$dispatch_detail_diff))) %>%
    mutate(category = "All") 

y <- data_orders %>%
    group_by(future_order, priority) %>%
    do(tidy(summary(.$dispatch_detail_diff))) %>%
    rename(category = priority)

z <- summary(data_orders$dispatch_detail_diff) %>%
    tidy() %>%
    mutate(category = "All",
           future_order = "All")

bind_rows(x, y, z) %>%
    select(type = future_order, category, everything()) %>%
    arrange(type, category) %>%
    knitr::kable(digits = 3, caption = "Summary measures for the time from dispatch to scheduled collection")

```

```{r, fig.cap="Quarter Hour of Task Dispatch. Most tasks dispatch on the 45-minute mark of the hour."}
df <- data_orders %>%
    filter(!is.na(Dispatched)) %>%
    mutate(dispatch_min = round_date(Dispatched, unit = "15 mins"),
           dispatch_min = minute(dispatch_min),
           dispatch_hour = round_date(Dispatched, unit = "hour"),
           dispatch_hour = hour(dispatch_hour))

plot_ly(df, x = ~factor(dispatch_min)) %>%
    add_histogram(split = ~interaction(priority, future_order))
```

```{r, fig.cap="Dispatch by hour of day"}
plot_ly(df, x = ~factor(dispatch_hour)) %>%
    add_histogram(split = ~interaction(priority, future_order))
```


# Future Orders

```{r, fig.cap="Distribution of time from task dispatch to scheduled collection by day of week"}
data_orders %>%
    filter(future_order == "Future") %>%
    plot_ly(x = ~day, y = ~dispatch_detail_diff, split = ~priority) %>%
    add_boxplot() %>%
    layout(xaxis = list(title = "Day of Week"), yaxis = list(title = "Time (Hours)"))
```

Day of week and hour of day are based on the scheduled collection time.

```{r, fig.cap="Distribution of time from task dispatch to scheduled collection by hour of day"}
data_orders %>%
    filter(future_order == "Future") %>%
    plot_ly(x = ~hour, y = ~dispatch_detail_diff, split = ~priority) %>%
    add_boxplot() %>%
    layout(xaxis = list(title = "Hour of Day", dtick = 2), yaxis = list(title = "Time (Hours)"))
```

```{r, fig.cap="Distribution of time from task order to dispatch"}
plot_ly(data_orders, x = ~order_dispatch_diff) %>%
    add_histogram(name = "All") %>%
    add_histogram(split = ~priority, visible = FALSE) %>%
    add_histogram(split = ~future_order, visible = FALSE) %>%
    layout(xaxis = list(title = "Time (Hours)"),
           updatemenus = list(list(
               y = 0.8, 
               buttons = list(
                   list(method = "restyle", args = list("visible", c(list(TRUE), as.list(rep(FALSE, times = 6)))), label = "All"),
                   list(method = "restyle", args = list("visible", c(list(FALSE), as.list(rep(TRUE, times = 4)), list(FALSE, FALSE))), label = "Priority"),
                   list(method = "restyle", args = list("visible", c(list(FALSE), as.list(rep(FALSE, times = 4)), list(TRUE, TRUE))), label = "Timing")
               ))
           ))
```

```{r, fig.cap="Distribution of time from task order to dispatch"}
plot_ly(data_orders, x = ~priority, y = ~order_dispatch_diff) %>%
    add_boxplot(name = "Overall") %>%
    add_boxplot(split = ~future_order) %>%
    layout(xaxis = list(title = "Collection Priority"), 
           yaxis = list(title = "Time (Hours)"))
```

# Control Charts

```{r, fig.cap="Time from dispatch until scheduled collection for Future Early AM orders"}
data_orders %>%
    filter(priority == "Early AM",
           future_order == "Future") %>%
    dmap_at("dispatch_detail_diff", ~ .x * 60) %>%
    qic(y = dispatch_detail_diff, data = ., chart = "i")
```

```{r, fig.cap="Time from dispatch until scheduled collection for Future Routine orders"}
data_orders %>%
    filter(priority == "Routine",
           future_order == "Future") %>%
    dmap_at("dispatch_detail_diff", ~ .x * 60) %>%
    qic(y = dispatch_detail_diff, data = ., chart = "i")
```

```{r, fig.cap="Time from dispatch until scheduled collection for Future Stat orders"}
data_orders %>%
    filter(priority == "Stat",
           future_order == "Future") %>%
    dmap_at("dispatch_detail_diff", ~ .x * 60) %>%
    qic(y = dispatch_detail_diff, data = ., chart = "i")
```

```{r, fig.cap="Time from dispatch until scheduled collection for Future Timed Study orders"}
data_orders %>%
    filter(priority == "Timed Study",
           future_order == "Future") %>%
    dmap_at("dispatch_detail_diff", ~ .x * 60) %>%
    qic(y = dispatch_detail_diff, data = ., chart = "i")
```

# Correlation Assessment

```{r, fig.cap="Time between order to scheduled collection and order to dispatch for future orders"}
data_orders %>%
    plot_ly(x = ~order_detail_diff, y = ~order_dispatch_diff, color = ~priority, split = ~future_order) %>%
    add_markers(marker = list(symbol = "circle-open")) %>%
    layout(xaxis = list(title = "Time from Order to Collection (hours)", range = c(-3, 48)), 
           yaxis = list(title = "Time from Order to Dispatch (hours)", range = c(-3, 48)))
```

There is a strong correlation between the time from order to scheduled collection and the time from order to dispatch.

```{r}
x <- data_orders %>%
    group_by(future_order) %>%
    do(tidy(cor.test(~ order_detail_diff + order_dispatch_diff, data = .))) %>%
    mutate(category = "All") 

y <- data_orders %>%
    filter(!(future_order == "Now" & priority == "Early AM")) %>%
    group_by(future_order, priority) %>%
    do(tidy(cor.test(~ order_detail_diff + order_dispatch_diff, data = .))) %>%
    rename(category = priority)

z <- data_orders %>%
    cor.test(~ order_detail_diff + order_dispatch_diff, data = .) %>%
    tidy() %>%
    mutate(category = "All",
           future_order = "All")

bind_rows(x, y, z) %>%
    select(type = future_order, category, everything()) %>%
    arrange(type, category) %>%
    knitr::kable(digits = 3, caption = "Correlation between the time from order to scheduled collection and the time from order to dispatch")
```

```{r, fig.cap="Time between order to scheduled collection and dispatch to scheduled collection"}
data_orders %>%
    plot_ly(x = ~order_detail_diff, y = ~dispatch_detail_diff, color = ~priority, split = ~future_order) %>%
    add_markers(marker = list(symbol = "circle-open")) %>%
    layout(xaxis = list(title = "Time from Order to Collection (hours)", range = c(-3, 48)), 
           yaxis = list(title = "Time from Dispatch to Collection (hours)", range = c(-3, 24)))
```

There is little correlation between the time from order to scheduled collection and time from dispatch to scheduled collection. This suggests that the dispatch time is triggered by the scheduled collection time, regardless of how far in advance the lab is ordered, with the exception of Stat orders which dispatch immediately.

```{r}
x <- data_orders %>%
    group_by(future_order) %>%
    do(tidy(cor.test(~ order_detail_diff + dispatch_detail_diff, data = .))) %>%
    mutate(category = "All") 

y <- data_orders %>%
    filter(!(future_order == "Now" & priority == "Early AM")) %>%
    group_by(future_order, priority) %>%
    do(tidy(cor.test(~ order_detail_diff + dispatch_detail_diff, data = .))) %>%
    rename(category = priority)

z <- data_orders %>%
    cor.test(~ order_detail_diff + dispatch_detail_diff, data = .) %>%
    tidy() %>%
    mutate(category = "All",
           future_order = "All")

bind_rows(x, y, z) %>%
    select(type = future_order, category, everything()) %>%
    arrange(type, category) %>%
    knitr::kable(digits = 3, caption = "Correlation between the time from order to scheduled collection and time from dispatch to scheduled collection")
```

```{r, fig.cap="Time between dispatch to scheduled collection and scheduled to actual collection"}
data_orders %>%
    plot_ly(x = ~dispatch_detail_diff, y = ~collect_detail_diff / 60, color = ~priority, split = ~future_order) %>%
    add_markers(marker = list(symbol = "circle-open")) %>%
    layout(xaxis = list(title = "Time from Dispatch to Collection (hours)", range = c(-3, 24)), 
           yaxis = list(title = "Time from Scheduled to Actual Collection (hours)", range = c(-12, 24)))
# x-tickvals = seq(0, 72, 6)
# y-tickvals = seq(-48, 120, 12)
```

There is no correlation between the amount of time prior to scheduled collection that the task dispatched and when the lab was actually collected.

```{r}
x <- data_orders %>%
    group_by(future_order) %>%
    do(tidy(cor.test(~ dispatch_detail_diff + collect_detail_diff, data = .))) %>%
    mutate(category = "All") 

y <- data_orders %>%
    filter(!(future_order == "Now" & priority == "Early AM")) %>%
    group_by(future_order, priority) %>%
    do(tidy(cor.test(~ dispatch_detail_diff + collect_detail_diff, data = .))) %>%
    rename(category = priority)

z <- data_orders %>%
    cor.test(~ dispatch_detail_diff + collect_detail_diff, data = .) %>%
    tidy() %>%
    mutate(category = "All",
           future_order = "All")

bind_rows(x, y, z) %>%
    select(type = future_order, category, everything()) %>%
    arrange(type, category) %>%
    knitr::kable(digits = 3, caption = "Correlation between the amount of time prior to scheduled collection that the task dispatched and when the lab was actually collected")
```

# Back-Dated Orders

```{r, fig.cap="Provider role that ordered back-dated levels"}
data_orders_backdated %>%
    dmap_at("action.provider.role", fct_infreq) %>%
    count(action.provider.role) %>%
    plot_ly() %>%
    add_bars(x = ~n, y = ~action.provider.role, orientation = "h") %>%
    layout(yaxis = list(title = ""), margin = list(l = 240))
```

```{r, fig.cap="Order type used for back-dated levels"}
data_orders_backdated %>%
    dmap_at("action.comm", fct_infreq) %>%
    count(action.comm) %>%
    plot_ly() %>%
    add_bars(y = ~n, x = ~action.comm) %>%
    layout(xaxis = list(title = "Order Type"), yaxis = list(title = "Count"))
```

# Now vs. Future Orders

```{r eval=FALSE}
# use this to identify specific individuals who placed the orders

library(edwr)
x <- data_orders %>%
    filter(priority %in% c("Routine", "Stat"),
           future_order == "Future",
           !early_am) 

y <- data_orders %>%
    filter(priority == "Timed Study",
           future_order == "Now",
           !early_am)

order_actions <- read_data("../data/raw", "action") %>%
    as.order_action() 

future_routine <- order_actions %>%
    semi_join(x, by = "order.id") %>%
    filter(action.type == "Order",
           action.provider != "SYSTEM") %>%
    select(order.id, action.provider, action.provider.role, action.comm) %>%
    count(action.provider)

now_timed <- order_actions %>%
    semi_join(y, by = "order.id") %>%
    filter(action.type == "Order",
           action.provider != "SYSTEM") %>%
    select(order.id, action.provider, action.provider.role, action.comm) %>%
    count(action.provider)

```

```{r, fig.cap="Orders for future labs by each order priority. Future labs are considered those > 1 hour from the time of ordering."}
x <- data_orders %>%
    # filter(!early_am) %>%
    dmap_at("future_order", ~ .x == "Future") %>%
    group_by(priority) %>%
    summarize_at("future_order", mean)

data_orders %>%
    # filter(!early_am) %>%
    plot_ly(x = ~priority, split = ~future_order) %>%
    add_histogram() %>%
    layout(xaxis = list(title = "Order Priority"))
```

Early AM orders are categorized as Routine. 

```{r, fig.cap="Provider role placing future orders using Routine or Stat"}
# who is placing future orders using routine / stat
data_orders %>%
    filter(priority %in% c("Routine", "Stat"),
           future_order == "Future") %>%
    dmap_at("action.provider.role", fct_infreq) %>%
    dmap_at("action.provider.role", fct_lump, n = 10) %>%
    count(action.provider.role) %>%
    plot_ly() %>%
    add_bars(x = ~n, y = ~action.provider.role, orientation = "h") %>%
    layout(xaxis = list(title = "Count"), yaxis = list(title = ""), margin = list(l = 240))
```

```{r, fig.cap="Provider role placing Now orders using Timed Study"}
# who is placing orders for now using timed study
data_orders %>%
    filter(priority == "Timed Study",
           future_order == "Now") %>%
    dmap_at("action.provider.role", fct_infreq) %>%
    dmap_at("action.provider.role", fct_lump, n = 10) %>%
    count(action.provider.role) %>%
    plot_ly() %>%
    add_bars(x = ~n, y = ~action.provider.role, orientation = "h") %>%
    layout(xaxis = list(title = "Count"), yaxis = list(title = ""), margin = list(l = 240))
```
