---
title: "Appropriate Vancomycin Level Collection in CVICU"
subtitle: "CS&E Project"
author: "Brian Gulbis, PharmD, BCPS"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_notebook:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(stringr)
library(lubridate)
library(tableone)
library(pander)

dirr::get_rds("../data/tidy")

hvi <- c("HH CVICU", "HH CVIMU", "HH HFIC", "HH HFIM", "HH 5HVI", "HH CCU", "HVI CIMU")

orders_levels <- full_join(levels, orders_valid, by = c("pie.id", "order.id")) %>%
    filter(event.unit %in% hvi | (is.na(event.unit) & order.unit %in% hvi),
           request == FALSE) %>%
    mutate(not_collected = is.na(Collected),
           location = coalesce(event.unit, order.unit),
           month = floor_date(detail.datetime, unit = "month"),
           day = wday(detail.datetime, label = TRUE, abbr = FALSE),
           hour = hour(detail.datetime)) %>%
    dmap_at(c("timely90", "timely60", "timely30"), ~ coalesce(.x, FALSE))

collected <- orders_levels %>%
    filter(!is.na(Collected)) %>%
    group_by(event.unit) %>%
    summarize(collected = n(),
              timely = sum(timely60, na.rm = TRUE),
              early = sum(early, na.rm = TRUE),
              late = sum(late, na.rm = TRUE)) 

not_done <- orders_levels %>%
    filter(is.na(Collected),
           is.na(lab.result)) %>%
    group_by(order.unit) %>%
    count %>%
    rename(not_done = n)

requests <- orders_requests %>%
    group_by(order.unit) %>%
    count %>%
    rename(requests = n) %>%
    filter(order.unit %in% hvi)

# percent_timely = timely / collected,
           
totals_units <- left_join(collected, not_done, by = c("event.unit" = "order.unit")) %>%
    left_join(requests, by = c("event.unit" = "order.unit")) %>%
    mutate(total = collected + not_done,
           percent_not_done = not_done / total,
           percent_appropriate = timely / total)

totals_hvi <- totals_units %>%
    ungroup() %>%
    summarize_at(vars(collected, timely, early, late, not_done, requests, total), funs(sum), na.rm = TRUE) %>%
    mutate(event.unit = "HVI",
           percent_not_done = not_done / total,
           percent_appropriate = timely / total) 

totals <- bind_rows(totals_units, totals_hvi) %>%
    select(event.unit, num_levels = total, everything()) %>%
    filter(!is.na(event.unit))

```

## Result Summary

```{r, fig.cap="Appropriately Collected Vancomycin Levels"}
df <- totals %>%
    filter(event.unit != "HVI")

hvi_avg <- totals$percent_appropriate[totals$event.unit == "HVI"]

ggplot(df, aes(x = event.unit, y = percent_appropriate)) +
    geom_bar(stat = "identity") +
    geom_hline(yintercept = hvi_avg, color = "blue") +
    ggtitle("Appropriately Collected Vancomycin Levels") +
    xlab("Unit") +
    ylab("Percent of Levels")
    # theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6))
```

* Blue line represents HVI average

```{r, fig.cap="Vancomycin Levels Ordered But Not Collected"}
df <- totals %>%
    filter(event.unit != "HVI")

hvi_avg <- totals$percent_not_done[totals$event.unit == "HVI"]

ggplot(df, aes(x = event.unit, y = percent_not_done)) +
    geom_bar(stat = "identity") +
    geom_hline(yintercept = hvi_avg, color = "blue") +
    ggtitle("Vancomycin Levels Ordered But Not Collected") +
    xlab("Unit") +
    ylab("Percent of Levels") 
```

```{r, fig.cap="Time Difference Between Scheduled and Actual Level Collection"}
ggplot(orders_levels, aes(x = location, y = collect_detail_diff)) +
    geom_boxplot() +
    # ggtitle("Time Difference Between Scheduled and Actual Level Collection") +
    coord_cartesian(ylim = c(-250, 250))
```

#### Raw Data

```{r, results='asis'}
vars <- c("not_collected", "timely90", "timely60", "timely30", "collect_detail_diff", "priority", "shift")
tbl_hvi <- CreateTableOne(vars, data = orders_levels)
tbl_units <- CreateTableOne(vars, "location", data = orders_levels)
ptbl_hvi <- print(tbl_hvi, printToggle = FALSE)
ptbl_units <- print(tbl_units, printToggle = FALSE)
rownames(ptbl_hvi) <- str_replace_all(rownames(ptbl_hvi), "   ", "- ")
total <- cbind(ptbl_hvi, ptbl_units)
knitr::kable(total[, 1:9], caption = "Raw Data for HVI and By Unit")
```

### Orders Placed by Vancomycin Request

```{r, results='asis'}
df <- orders_requests %>%
    filter(order.unit %in% hvi) %>%
    mutate(month = floor_date(detail.datetime, unit = "month"),
           day = wday(detail.datetime, label = TRUE, abbr = FALSE),
           hour = hour(detail.datetime)) 

vars <- c("shift")
tbl_hvi <- CreateTableOne(vars, data = df)
tbl_units <- CreateTableOne(vars, "order.unit", data = df)
ptbl_hvi <- print(tbl_hvi, printToggle = FALSE)
ptbl_units <- print(tbl_units, printToggle = FALSE)
rownames(ptbl_hvi) <- str_replace_all(rownames(ptbl_hvi), "   ", "- ")
total <- cbind(ptbl_hvi, ptbl_units)
knitr::kable(total[, 1:9], caption = "Orders by Vancomycin Request")
```

## Exploration of CVICU Levels
```{r}
cvicu <- orders_levels %>%
    filter(location == "HH CVICU") %>%
    rename(appropriate = timely60)
```

### By Shift
```{r}
vars <- c("appropriate", "not_collected") 
tbl <- CreateTableOne(vars, "shift", cvicu)
ptbl <- print(tbl, printToggle = FALSE)
knitr::kable(ptbl[, 1:3], caption = "Vancomycin Levels in CVICU by Nursing Shift")
```

### By Month

```{r}
vars <- c("appropriate", "not_collected", "shift") 
tbl <- CreateTableOne(vars, "month", cvicu)
ptbl <- print(tbl, printToggle = FALSE)
knitr::kable(ptbl[, c(2:7, 9)], caption = "Vancomycin Levels in CVICU by Month")
```

### By Day of Week

```{r}
vars <- c("appropriate", "not_collected", "shift") 
tbl <- CreateTableOne(vars, "day", cvicu)
ptbl <- print(tbl, printToggle = FALSE)
knitr::kable(ptbl[, 1:8], caption = "Vancomycin Levels in CVICU by Day of Week")
```

### By Hour of Day

```{r, fig.cap="Appropriately vs. Inappriately Collected Levels by Hour of Day"}
ggplot(cvicu, aes(x = hour)) +
    geom_bar(aes(fill = appropriate)) +
    ggtitle("Appropriately vs. Inappriately Collected Levels by Hour of Day")
```


```{r, fig.cap="Collected vs. Un-Collected Levels by Hour of Day"}
ggplot(cvicu, aes(x = hour)) +
    geom_bar(aes(fill = not_collected)) +
    ggtitle("Collected vs. Un-Collected Levels by Hour of Day")
```


```{r}
ggplot(orders_levels, aes(x = hour)) +
    geom_bar(aes(fill = not_collected)) +
    ggtitle("Collected vs. Un-Collected Levels by Hour of Day") +
    facet_wrap(~ location)
```

### Timing

Distribution of difference between desired draw time and actual draw time.

```{r, fig.cap="Distribution of Time Difference Between Desired and Actual Collection"}
mn <- mean(cvicu$collect_detail_diff, na.rm = TRUE)
std <- sd(cvicu$collect_detail_diff, na.rm = TRUE)

ggplot(cvicu, aes(x = collect_detail_diff)) +
    geom_density() +
    geom_vline(xintercept = mn, color = "blue") +
    geom_vline(xintercept = mn + c(1, -1) * std, color = "red") +
    geom_vline(xintercept = c(-60, 60), color = "darkgreen", linetype = "dashed") +
    ggtitle("Distribution of Time Difference")
# coord_cartesian(xlim = c(-200, 200))
```

* Blue line represents mean
* Red lines represent standard deviation
* Green dashed lines represent 60 minutes

#### Remove extreme outliers

```{r, fig.cap="Distribution of Time Difference Without Outliers"}
q <- quantile(cvicu$collect_detail_diff, na.rm = TRUE, probs = c(0.25, 0.5, 0.75))
iqr <- IQR(cvicu$collect_detail_diff, na.rm = TRUE)

no_outliers <- cvicu %>%
    filter(!is.na(collect_detail_diff),
           collect_detail_diff >= q[1] - (iqr * 3),
           collect_detail_diff <= (iqr * 3) + q[3])

mn <- mean(no_outliers$collect_detail_diff, na.rm = TRUE)
std <- sd(no_outliers$collect_detail_diff, na.rm = TRUE)

ggplot(no_outliers, aes(x = collect_detail_diff)) +
    geom_density() +
    geom_vline(xintercept = mn, color = "blue") +
    geom_vline(xintercept = mn + c(1, -1) * std, color = "red") +
    geom_vline(xintercept = c(-60, 60), color = "darkgreen", linetype = "dashed") +
    ggtitle("Distribution of Time Difference Without Outliers")
```

Relationship between how far in advance the level was ordered, and how close to desired draw time the level was actually drawn.

```{r, fig.cap="Time from order to draw vs. time to actual draw"}
cvicu %>%
    filter(!is.na(sched_diff),
           !is.na(collect_detail_diff)) %>%
    ggplot(aes(x = as.numeric(sched_diff), y = as.numeric(collect_detail_diff))) +
    geom_point(aes(color = shift), alpha = 0.6) +
    scale_x_continuous(breaks = seq(0, 96, 12)) +
    ggtitle("Time from order to draw vs. time to actual draw") +
    xlab("Ordering to desired draw (hours)") +
    ylab("Desired draw to actual draw (minutes)")
```

### Time Differences

Time (in minutes) between when levels was scheduled to be drawn and when it was actually drawn.

#### By Shift
```{r, fig.cap="Time Difference Between Scheduled and Actual Level Collection"}
ggplot(cvicu, aes(x = shift, y = as.numeric(collect_detail_diff))) +
    geom_boxplot() +
    ggtitle("Time Difference Between Scheduled and Actual Level Collection") +
    coord_cartesian(ylim = c(-250, 250))
```


```{r}
t.test(collect_detail_diff ~ shift, cvicu)
```

## Notes

* There are automatic vancomycin requests created by a Discern Advisor
    - Unclear where this is generated from
    - Request detail was usually several days prior to the order date
* Levels ordered as Early AM generate a child order with a scheduled time of 0300 the following day
    - The parent Early AM order was removed from this analysis due to an invalid scheduled collection time
